---
title: "Regression: Full cohort"
author: "S. Mason"
date: "10/23/2018"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library (ggplot2)
library(survival)
library(survminer)
library(MASS)
library(psych)
library(gmodels)
```
```{r load data, message=FALSE, warnings=FALSE, echo=FALSE}
#Load outcomes data
outcomes <- read_excel("~/OneDrive - Partners HealthCare/research/ROCI/CXR/data/ROCI_101918.xlsx", sheet = "Outcomes", col_types = c("numeric", "skip", "numeric", "numeric", "numeric", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "date", "date", "numeric", "numeric", "date", "date",  "numeric", "date", "date", "numeric", "text", "numeric", "text", "numeric", "numeric", "date", "numeric", "numeric", "text", "numeric", "text", "text"))
#Load imaging/score data
images <- read_excel("~/OneDrive - Partners HealthCare/research/ROCI/CXR/data/ROCI_101918.xlsx", sheet = "Imaging", col_types = c("numeric", "text", "date", "numeric", "date", "numeric", "numeric", "numeric", "date", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "skip"))
```

<h3>Analysis of the non-control cohort who have a CXR from MICU admit (n=559)</h3>
**TLDR**:
<br> - Total CXR score IS predictive of in-hospital mortality, even when adjusted for APACHE score and immunosuppression
<br> - Total CXR score IS predictive of 28-day mortality, even when adjusted for APACHE score and immunosuppression
<br> - Total CXR score IS predictive of 60-day mortality, even when adjusted for APACHE score and immunosuppression
<br> - Total CXR score IS predictive of time-to-death ina Cox proportional hazards model
<br> - Total CXR score IS predictive of ICU LOS and duration of MV
<br>
```{r filter data}
#Creates one dataset of outcomes and scores for subjects with a CXR from MICU admit
admit <- filter(images, Final_Dx != "Control", admit_cxr == 1) %>% select(Subject_ID, cxr_date, Quad1, Quad2, Quad3, Quad4, Total)
imgs <- filter(outcomes, Final_Dx != "Control", imgs == 1) %>% select (Subject_ID, Male, Age, BMI, non_white, Immsupp, APACHE, SOFA, days_MV, ICU_admit, ICU_days, Hosp_admit, Hosp_days, Final_Dx, DC_coded, hosp_death, date_death, mort28d, mort60d) %>% merge(admit, by="Subject_ID")

# 207/559 subjects are immunosuppressed
# 286/559 subjects have a BMI recorded
```
```{r more set-up, echo=FALSE}
#Numerics to factor
imgs$Male <- as.factor(imgs$Male)
imgs$non_white <- as.factor(imgs$non_white)
imgs$Immsupp <- as.factor(imgs$Immsupp)
imgs$hosp_death <- as.factor(imgs$hosp_death)
imgs$mort28d <- as.factor(imgs$mort28d)
imgs$mort60d <- as.factor(imgs$mort60d)
```

<br>**In-hospital mortality**: With the exception of age and gender, all of the covariates are significant predictors of in hospital death and none of them appear to be colinear or confounding. Notably, total CXR score remains significant even when adjusting for APACHE score and immunosuppression status. 
```{r logit in-hospital mortality}
#In-hospital mortality
hosp_mort <- glm(hosp_death ~ Total+Male+Age+Immsupp+APACHE+non_white, family=binomial(link='logit'), data=imgs)
summary(hosp_mort)

#Table of univariable analysis
univariable <- data_frame(var = c("total", "Age", "non_white", "immsupp", "APACHEII", "Male"), beta_tot = c(.1221, .12199, .11805, .10926, .11156, .12221), std_err_tot = c(.0234, .0234, .02378, .02454, .02448, .0234), pval_tot = c("<<.01", "<<.01", "<<.01", "<<.01", "<<.01", "<<.01"), AIC = c(614.3, 616.28, 602.97, 569.72, 583.11, 616.3), pval_var = c(NA, .907, .0008, "<<.01", "<<.01", .977), comment = c("predictor", "not sig", "predictor", "predictor", "predictor", "not sig"))
univariable
```
<br>**28-day mortality**: Age and gender are no longer predictive, but Total CXR score remains significant even when Immunosuppression and APACHE are adjusted for. 

```{r logit 28d mortality}
one_month <- glm(mort28d ~ Total+Male+Age+Immsupp+APACHE+non_white, family=binomial(link='logit'), data=imgs)
summary(one_month)
```
<br>**60-day mortality**: Age and gender are no longer predictive, but Total CXR score remains significant even when Immunosuppression and APACHE are adjusted for. 

```{r logit 60d mortality}
two_month <- glm(mort60d ~ Total+Male+Age+Immsupp+APACHE+non_white, family=binomial(link='logit'), data=imgs)
summary(two_month)
```
<br>**Time to death/Cox model**:  The total CXR score is the only predictive covariate in a cox proportional hazards model. 
```{r cox proportional hazards}
refdate <- as.Date("2018-09-15")
#Create variables with survival (in days) and a binary indicating censored status
imgs <- mutate(imgs, survival = ifelse(is.na(date_death) == TRUE, as.numeric(difftime(refdate, ICU_admit, units='days')), as.numeric(difftime(date_death, ICU_admit, units='days')))) %>% mutate(censor = ifelse(is.na(date_death) == TRUE, 0, 1))

#Cox regression model
cox_all <- coxph(Surv(survival, censor) ~ Male+Age+non_white+APACHE+Immsupp+Total, data=imgs)
summary(cox_all)
```
**Cox by quartile**: 
```{r quartiles}
imgs <- imgs %>% mutate(quartile = ntile(Total,4))
cox_quart <- coxph(Surv(survival, censor) ~ strata(quartile), data=imgs)
summary(cox_quart)
ggsurvplot(survfit(cox_quart), data=imgs, risk.table=TRUE, ggtheme=theme_minimal())
```
<br>**Other outcome measures**: Total (along with APACHE, age, and race) are predictive of both ICU LOS and duration of MV (in a simple linear regression model with a non-transformed outcome).
```{r other outcomes}
#Length of ICU stay
ICUdays <- glm(ICU_days ~ Total+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(ICUdays)

#Duration of mechanical ventilation
mvdays <- glm(days_MV ~ Total+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(ICUdays)
```

<br>**Other outcome measures**: Total (along with APACHE, age, and race) are predictive of both ICU LOS and duration of MV (in a simple linear regression model with a non-transformed outcome) - Using negative binomial regression.
```{r other outcomes}
imgs$quartile_f <- as.factor(imgs$quartile)
contrasts(imgs$quartile_f) <- contr.treatment(4, base=1)

#Length of ICU stay
ICUdays_nb <- glm.nb(ICU_days ~ Total+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(ICUdays_nb)
cbind(IRR = exp(coef(ICUdays_nb)), 
      exp(confint(ICUdays_nb)))[2,]
### by quartile
ICUdays_nb_quart <- glm.nb(ICU_days ~ quartile_f+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(ICUdays_nb_quart)
cbind(IRR = exp(coef(ICUdays_nb_quart)), 
      exp(confint(ICUdays_nb_quart)))[2:4,]

#Duration of mechanical ventilation
mvdays_nb <- glm.nb(days_MV ~ Total+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(mvdays_nb)
cbind(IRR = exp(coef(mvdays_nb)), 
      exp(confint(mvdays_nb)))[2,]
### by quartile
mvdays_nb_quart <- glm.nb(days_MV ~ quartile_f+APACHE+Age+non_white+Male+Immsupp, data=imgs)
summary(mvdays_nb_quart)
cbind(IRR = exp(coef(mvdays_nb_quart)), 
      exp(confint(mvdays_nb_quart)))[2:4,]

#Duration of mechanical ventilation - subset who were intubated
imgs_MV <- imgs %>% filter(days_MV > 0) #subset
mvdays_nb_mvsub <- glm.nb(days_MV ~ Total+APACHE+Age+non_white+Male+Immsupp, data=imgs_MV)
summary(mvdays_nb_mvsub)
cbind(IRR = exp(coef(mvdays_nb_mvsub)), 
      exp(confint(mvdays_nb_mvsub)))[2,]
### by quartile
mvdays_nb_mvsub_quart <- glm.nb(days_MV ~ quartile_f+APACHE+Age+non_white+Male+Immsupp, data=imgs_MV)
summary(mvdays_nb_mvsub_quart)
cbind(IRR = exp(coef(mvdays_nb_mvsub_quart)), 
      exp(confint(mvdays_nb_mvsub_quart)))[2:4,]
```

<br>**Table 1 measures:**
```{r table 1}
#Note: there are more elegant solutions... This is just quick and dirty.

# numeric values
imgs_num <- imgs %>% select_if(is.numeric)
describe(imgs_num)

#categorical - percent male
CrossTable(imgs$Male)

#categorical - percent immunosuppressed
CrossTable(imgs$Immsupp)

#categorical - percent non-white
CrossTable(imgs$non_white)

#categorical - percent non-whitein hospital mortality)
```

